Mass and occupation fraction of dark matter halos hosting Lyman-$\alpha$ emitters at $z\sim 3$

### 2013-08-30 paper/ms.tex
new introduction done

\title[Halo mass and occupation fraction for LAEs at
  $z=3.1$]{Mass and occupation
  fraction of dark matter halos hosting Lyman-$\alpha$ emitters
  at $z\sim 3$}    
\author[~J.~E. Forero-Romero and ~J.~E Mejia-Restreo]{
\parbox[t]{\textwidth}{\raggedright 
  Jaime E. Forero-Romero$^{1}$ and
  Julian E. Mej\'ia-Restrepo$^{2}$ 
}
\vspace*{6pt}\\
$^{1}$ Departamento de F\'{i}sica, Universidad de los Andes, Cra. 1
No. 18A-10, Edificio Ip, Bogot\'a, Colombia \\
$^{2}$ Departamento de Astronom\'{i}a, Universidad de Chile, Camino el
Observatorio 1515, Santiago, Chile}

@@

\begin{abstract}
%
We derive constraints on the mass and occupation fraction of dark
matter halos hosting \ly Emitting galaxies (LAEs) at a redshift of
$z=3.1$ by matching the number density and the angular
correlation function between mock and observed fields. We explicitly
take into account the cosmic variance on the typical observed field size by
constructing mock fields from a large cosmological N-body
simulation matching the geometries of observed fields. To populate the
halos in the simulation we use a model where a dark matter halo with
mass in the range $M_{\rm   min}<M_{\rm h}<M_{\rm max}$ can only host
one LAE with a probability $f_{\rm occ}$. This main motivation to
build such a model is to avoid the large uncertainties in the estimation of
\ly luminosities. We find that the clustering and number density information are
insufficient to impose a tight constraint on the escape fraction. On
the other hand the minium mass and maximum mass are tightly
constrained to be $M_{\rm max}<10^{12}$\hMsun and
$10^{10}\hMsun\leq M_{\rm min}\leq 10^{11.5}\hMsun$.  We also find
that all models have a narrow mass range $\Delta M \equiv
\log_{10}M_{\rm max} - \log_{10} M_{\rm min}$ smaller than 1.0 dex,
with the vast majority in the range $\Delta M<0.5$ dex. All previous
constraints on the mass range and occupation fraction are a sub-set of
the models we report in this paper. This represents an unconclusive response as
to which is the occupation fraction for LAEs at $z=3.1$. A more
precise answer requires physical models for the connection between
dark matter, star formation and the escape of \ly radiation. That kind
of modeling should also provide a plausible explanation for the narrow
 ange in mass $\Delta M<1.0$ dex derived here for the halos hosting
 LAEs at $z=3.1$.

@@

Lyman-$\alpha$ emitting galaxies (LAEs) have become in the last decade
a  central topic in studies of structure formation in the Universe. They 
are helpful in a diverse range of fields. LAEs can be used as probes
of reionization \citep{Dijkstra11}, tracers of large scale structure
\citep{Koehler2007},  signposts for low metallicity stellar
populations, markers of the galaxy formation process through
cosmic history \citep{ForeroRomero2012} and tracers of active star
formation  \citep{Guaita2013}

At the same time, theoretical and observational developments have
contributed to the emergence of a paradigm to describe structure
formation in a cosmological context. In this context it is considered
that dominant matter content of the Universe is to be found in dark
matter (DM), whereby each galaxy is hosted by larger dark matter structure
known as a halo. \citep{Peebles1980,SpringelNature05}.

Most models of galaxy formation find that halo mass can be
used to predict galaxy properties such as the stellar mass and
star formation rate \citep{Behroozi2012}. Processes that regulate the
star formation cycle are also though to be strongly dependent on its
mass.  For these reasons, finding the typical dark matter halo mass
hosting LAEs represents a significant step forward to understand the
nature of this galaxy population in the context of Lambda Cold Dark
Matter ($\Lambda$CDM) paradigm.

Some theoretical approaches to this problem have been based on
ab-initio approach. Starting from the DM halo population, the
corresponding intrinsic star formation properties are inferred
together with other statistics such as the luminosity function, the
correlation function and the equivalent width distributions. Such
modelling has been implemented from analytic considerations,
semi-analytic models  and  full N-body hidrodynamical simulations
\citep{Laursen2007, Dayal2009, ForeroRomero2011, Yajima2012, Orsi2012,
  Soler2012}.

In addition to the uncertainties in the astrophysical processes
describing star formation in galactic populations, is the calculation
of the fraction of \ly photons that escape the galaxy to the observer
is another debated step. Given the resonant nature of the \ly line,
the radiative transfer of \ly is sensitive to the density,
temperature, topology and kinematics of the neutral Hydrogen in the
interstellar medium
(ISM). \citep{Neufeld1991,ForeroRomero2011,Dijkstra2012,Laursen2013,Orsi2012}.

This problem can be tackled with Monte Carlo simulations for the
radiative transfer, however the complexity of the relevant physical
processes makes it difficult to achieve a robust consensus on what is
the theoretical expected value for the \ly escape fraction
at high redshift.

Another approach to infer the typical DM halo mass of halos hosting
LAEs is based on the spatial information. This uses the fact that in cold
dark matter cosmologies the spatial clustering of galaxies on large
scales is entirely dictated by the halo distribution
\citep{Colberg00}, which in turn has a strong dependence on halo
mass. Using measurements of the angular correlation function of LAEs,
observers have put constraints on the typical mass and occupation
fraction of the putative halos hosting these galaxies
\citep{Hayashino2004,Gawiser07,Nilsson2007,Ouchi2010}. In this
studies the observations were done on single fields of $\sim 1$
deg$^{2}$ and the conclusions derived on the halo host mass from
clustering information do not delve to deeply into the impact of
cosmic variance.

@@

In this \documentname to study the impact of cosmic variance,
we implement a method to populate the DM halos in cosmological
simulations with LAEs. This approach bypasses all the uncertainty
involved in the estimation of star formation rates and \ly escape
fraction. The model only considers whether a DM can host a
detectable LAE or not. It does not predict a \ly  luminosity. Once the
mock catalogs are constructed  we proceed with a direct comparison
against the statistics derived from observations. This allows us 
to find the preferred mass range of DM halos hosting these galaxies
and its occupation fraction.

>>>>

\title[Halo mass and occupation fraction of LAEs at$z=3.1$]{Mass and
  occupation fraction of dark matter halos hosting Lyman-$\alpha$
  emitters at $z\sim 3$}      
\author[~J.~E. Forero-Romero and ~J.~E. Mejia-Restreo]{
\parbox[t]{\textwidth}{\raggedright 
  Jaime E. Forero-Romero$^{1}$ and
  Julian E. Mej\'ia-Restrepo$^{2}$ 
}
\vspace*{6pt}\\
$^{1}$ Departamento de F\'{i}sica, Universidad de los Andes, Cra. 1
No. 18A-10, Edificio Ip, Bogot\'a, Colombia \\
$^{2}$ Departamento de Astronom\'{i}a, Universidad de Chile, Camino el
Observatorio 1515, Santiago, Chile}

@@

\begin{abstract}
%
We derive constraints on the mass and occupation fraction of dark
matter halos hosting \ly Emitting galaxies (LAEs) at a redshift of
$z=3.1$ by matching the number density and the angular
correlation function between mock and observed fields. We explicitly
take into account the cosmic variance on the typical observed field size by
constructing mock fields from a large cosmological N-body
simulation matching the geometries of observed fields. To populate the
halos in the simulation we use a model where a dark matter halo with
mass in the range $M_{\rm   min}<M_{\rm h}<M_{\rm max}$ can only host
one detectable LAE with a probability $f_{\rm occ}$. The main motivation to
build such a model is to avoid the large uncertainties in the estimation of
\ly luminosities. We find that the clustering and number density information are
insufficient to impose a tight constraint on the escape fraction. On
the other hand the minium mass and maximum mass are tightly
constrained to be $M_{\rm max}<10^{12}$\hMsun and
$10^{10}\hMsun\leq M_{\rm min}\leq 10^{11.5}\hMsun$.  We also find
that all models have a narrow mass range $\Delta M \equiv
\log_{10}M_{\rm max} - \log_{10} M_{\rm min}$ smaller than $1.0$ dex,
with the vast majority in the range $\Delta M<0.5$ dex. Already
published constraints on the mass range and occupation fraction are a sub-set of
the models we report in this \documentname. Our results present an
inconclusive response as to which is the occupation fraction for LAEs
at $z=3.1$. A precise answer requires robust physical models for the
connection between dark matter, star formation and the escape of \ly
radiation. That kind of modelling should also be able to provide a
plausible explanation for the narrow ange in mass $\Delta M<1.0$ dex
derived here for the halos hosting LAEs at $z=3.1$.

@@

Lyman-$\alpha$ emitting galaxies (LAEs) have become in the last decade
a  central topic in studies of structure formation in the Universe. They 
are helpful in a diverse range of fields. LAEs can be used as probes
of reionization \citep{Dijkstra11}, tracers of large scale structure
\citep{Koehler2007},  signposts for low metallicity stellar
populations, markers of the galaxy formation process at high redshift
\citep{Dayal2009,ForeroRomero2012} and tracers of active star formation
\citep{Guaita2013}

At the same time, theoretical and observational developments have
contributed to the emergence of a paradigm to describe structure
formation in a cosmological context. In this context it is considered
that the dominant matter content of the Universe is to be found in dark
matter (DM), whereby each galaxy is hosted by larger dark matter structure
known as a halo. \citep{Peebles1980,SpringelNature05}.

Most models of galaxy formation find that halo mass can be
used to predict galaxy properties such as the stellar mass and
star formation rate \citep{Behroozi2012}. Processes that regulate the
star formation cycle are also though to be strongly dependent on its
halo mass.  For these reasons, finding the typical dark matter halo mass
hosting LAEs represents a significant step forward to understand the
nature of this galaxy population in the context of Lambda Cold Dark
Matter ($\Lambda$CDM) paradigm.

Some theoretical approaches to this problem have been based on
ab-initio approach. Starting from the DM halo population, the
corresponding intrinsic star formation properties are inferred
together with other statistics such as the luminosity function, the
correlation function and the equivalent width distributions. Such
modelling has been implemented from analytic considerations,
semi-analytic models \citep{Garel2012, Orsi2012}and  full N-body
hydrodynamical simulations \citep{Laursen2007, Dayal2009,
  ForeroRomero2011, Yajima2012, Soler2012}.

In addition to the uncertainties in the astrophysical processes
describing star formation in galactic populations,the calculation
of the fraction of \ly photons that escape the galaxy to the observer
is another debated step. Given the resonant nature of the \ly line,
the radiative transfer of \ly is sensitive to the density,
temperature, topology and kinematics of the neutral Hydrogen in the
interstellar medium
(ISM). This problem can be tackled with Monte Carlo simulations for the
radiative transfer, however the complexity of the relevant physical
processes makes it difficult to achieve a robust consensus on what is
the theoretical expected value for the \ly escape fraction
at high redshift
\citep{Neufeld1991,ForeroRomero2011,Dijkstra2012,Laursen2013,Orsi2012}.

Another approach to infer the typical DM halo mass of halos hosting
LAEs is based on the clustering information. This uses the fact that in cold
dark matter cosmologies the spatial clustering of galaxies on large
scales is entirely dictated by the halo distribution
\citep{Colberg00}, which in turn has a strong dependence on halo
mass. Using measurements of the angular correlation function of LAEs,
observers have put constraints on the typical mass and occupation
fraction of the putative halos hosting these galaxies
\citep{Hayashino2004,Gawiser07,Nilsson2007,Ouchi2010}. In this
studies the observations were done on single fields of $\sim 1$
deg$^{2}$ and the conclusions derived on the halo host mass from
clustering information do not delve to deeply into the impact of
cosmic variance.

@@

In this \documentname we implement a method to populate the DM halos
in cosmological simulations with LAEs that bypasses all the
uncertainty involved in the estimation of star formation rates and \ly escape
fraction. The model only considers whether a DM can host a
detectable LAE or not without predicting a \ly  luminosity. Once the
mock catalogs are constructed  we proceed with a direct comparison
against the statistics derived from observations. This allows us 
to find the preferred mass range of DM halos hosting these galaxies
and its occupation fraction.
